---
title: "V1 Statistical Analysis"
author: "Quentin Schorpp"
date: "Wednesday, September 17, 2014"
output: html_document
---

This is the Analysis of a microcosm experiment performed at the Th?nen Institute of Biodiversity, Braunschweig. The Experiment aimed to investigate the influence of soil fauna groups Lumbricidae (*L. terrestris*) and Collembola (*F. candida*) on the microbial activity involved in gaseous nitrogen emissions. The data consists of three different qualities of measurement. First emissions of N[2]O and CO[2] were measured continously during the whole duration of the experiment in intervals lasting several days. That data is in units of ppb or ppm. Second at four dates the microcosms were incubated with artificial atmosphere and 15N isotopic signatures of N2O and N2 were measured. Third at one particular date samples for isotopomere measurements were taken.

  
# 1. Eathworm Biomass
Earthworm biomass was measured in g after earthworm guts were voided. Mesurements took place before and after the experiment.
During the Experiment four Individual of L. terrestris died in Microcosms with the interaction Treatment in the 15N experimental part / approach.


# N2O and CO2 gas fluxes
Measurements from continous flow samples.
***********************************************************************************************************************************

# Analysis
```{r}
# data = n2o.cum

n2o.index$Lt <- revalue(n2o.index$treat, c(Lt=1, int=1, Fc=0, C=0))
n2o.index$Fc <- revalue(n2o.index$treat, c(Lt=0, int=1, Fc=1, C=0))

n2o.cum <- as.data.frame(cbind(n2o.index[c(1:56),][-c(8,22),], n2o.cum))
colnames(n2o.cum)[21] <- "nc"
colnames(n2o.cum)[4] <- "exp"

n2o.cum <- droplevels(n2o.cum)

### Experiment 1

# multimodel averaging
x <- c("Lt","Fc","exp","soil")
nc.0 <- lm(nc~Lt*Fc*soil, n2o.cum[!(n2o.cum$exp=="exp2"),])
#t(combn(x,3))
nc.1 <- lm(nc~Lt*Fc, n2o.cum[!(n2o.cum$exp=="exp2"),])
nc.2 <- lm(nc~Lt*soil, n2o.cum[!(n2o.cum$exp=="exp2"),])
nc.3 <- lm(nc~Fc*soil, n2o.cum[!(n2o.cum$exp=="exp2"),])
#t(combn(x,2))
nc.4 <- lm(nc~Lt, n2o.cum[!(n2o.cum$exp=="exp2"),])
nc.5 <- lm(nc~Fc, n2o.cum[!(n2o.cum$exp=="exp2"),])
nc.6 <- lm(nc~soil, n2o.cum[!(n2o.cum$exp=="exp2"),])

nc.7 <- lm(nc~1, n2o.cum[!(n2o.cum$exp=="exp2"),])

AICctab(nc.0,  nc.1,  nc.2,  nc.3,  nc.4,  nc.5,  nc.6,  nc.7)

#as.character(rep("nc",c(1:15)))
#x <- factor(paste("nc",".",c(0:15),",",sep=""), levels=)
#print(x)
#x <- factor(paste("nc",".",c(0:15)," <- ",sep=""), levels=)
#as.data.frame(x)

summary.aov(nc.2)
par(mfrow=c(2,2))
plot(nc.2)
par(mfrow=c(1,1))
boxcox(nc.2)
#locator(1)

nc.2 <- update(nc.2,I(1/nc^(1/5))~.)
summary.aov(nc.2)
par(mfrow=c(2,2))
plot(nc.2)

nc.aov2 <- aov(nc~Lt*soil, n2o.cum[!(n2o.cum$exp=="exp2"),])
nc.aov2 <- update(nc.aov2,I(1/nc^(1/5))~.)
TukeyHSD(nc.aov2)
nc.int <- with(n2o.cum[!(n2o.cum$exp=="exp2"),], interaction(soil,Lt)) # Double interaction factor
nc.aov3 <- update(nc.aov2,.~nc.int)
HSD.test(nc.aov3, "nc.int", group=TRUE, console=TRUE)

txt.nc = expression(paste("Nitrous Oxide ", N[2],"O [?g ",N[2],"O-N ",kg^-1,"]"))
ggplot(n2o.cum[!(n2o.cum$exp=="exp2"),], aes(x=Lt, y=nc), family="Arial") + 
  stat_boxplot(geom="errorbar", coef=1.5, lwd=0.2) + 
  geom_boxplot(fill="grey", lwd=0.2, outlier.size=0.4, outlier.shape=21) +  
  facet_grid( .~ soil , scales="free", space="free",labeller=label_parsed) + 
  ylab(txt.nc) + 
  xlab("L. terrestris") + 
  scale_x_discrete(labels=c("absent", "present")) +
  ggtitle("Experiment 1") +
  mytheme


nc.int <- with(n2o.cum[!(n2o.cum$exp=="exp2"),], interaction(soil,Lt)) # Double interaction factor
nc.aov3 <- update(nc.aov2,.~nc.int)
HSD.test(nc.aov3, "nc.int", group=TRUE, console=TRUE)

with(n2o.cum[!(n2o.cum$exp=="exp1"),], list(round(tapply(nc, interaction(treat,soil), mean),0),
                                            round(tapply(nc, interaction(treat,soil), se),0)))
                                            
                                            
with(n2o.cum[!(n2o.cum$exp=="exp2"),], list(round(tapply(nc, interaction(Lt,soil), mean),0),
                   round(tapply(nc, interaction(Lt,soil), se),0)))


### Experiment 1

# multimodel averaging
x <- c("Lt","Fc","exp","soil")
nc.0 <- lm(nc~Lt*Fc*soil, n2o.cum[!(n2o.cum$exp=="exp1"),])
#t(combn(x,3))
nc.1 <- lm(nc~Lt*Fc, n2o.cum[!(n2o.cum$exp=="exp1"),])
nc.2 <- lm(nc~Lt*soil, n2o.cum[!(n2o.cum$exp=="exp1"),])
nc.3 <- lm(nc~Fc*soil, n2o.cum[!(n2o.cum$exp=="exp1"),])
#t(combn(x,2))
nc.4 <- lm(nc~Lt, n2o.cum[!(n2o.cum$exp=="exp1"),])
nc.5 <- lm(nc~Fc, n2o.cum[!(n2o.cum$exp=="exp1"),])
nc.6 <- lm(nc~soil, n2o.cum[!(n2o.cum$exp=="exp1"),])

nc.7 <- lm(nc~1, n2o.cum[!(n2o.cum$exp=="exp1"),])

AICctab(nc.0,  nc.1,  nc.2,  nc.3,  nc.4,  nc.5,  nc.6,  nc.7)

#as.character(rep("nc",c(1:15)))
#x <- factor(paste("nc",".",c(0:15),",",sep=""), levels=)
#print(x)
#x <- factor(paste("nc",".",c(0:15)," <- ",sep=""), levels=)
#as.data.frame(x)

summary.aov(nc.2)
summary.aov(nc.0)
par(mfrow=c(2,2))
plot(nc.2)
par(mfrow=c(1,1))
boxcox(nc.2)
# locator(1)
fligner.test(nc~interaction(Lt,soil), n2o.cum[!(n2o.cum$exp=="exp1"),])


nc.aov2 <- aov(nc~Lt*soil, n2o.cum[!(n2o.cum$exp=="exp1"),])
TukeyHSD(nc.aov2)
nc.int <- with(n2o.cum[!(n2o.cum$exp=="exp1"),], interaction(soil,Lt)) # Double interaction factor
nc.aov3 <- update(nc.aov2,.~nc.int)
HSD.test(nc.aov3, "nc.int", group=TRUE, console=TRUE)

txt.nc = expression(paste("Nitrous Oxide ", N[2],"O [?g ",N[2],"O-N ",kg^-1,"]"))
ggplot(n2o.cum[!(n2o.cum$exp=="exp1"),], aes(x=Lt, y=nc), family="Arial") + 
  stat_boxplot(geom="errorbar", coef=1.5, lwd=0.2) + 
  geom_boxplot(fill="grey", lwd=0.2, outlier.size=0.4, outlier.shape=21) +  
  facet_grid( .~ soil , scales="free", space="free",labeller=label_parsed) + 
  ylab(txt.nc) + 
  xlab("L. terrestris") + 
  scale_x_discrete(labels=c("absent", "present")) +
  ggtitle("Experiment 2") +
  mytheme


nc.int <- with(n2o.cum[!(n2o.cum$exp=="exp1"),], interaction(soil,Lt,Fc)) # Double interaction factor
nc.aov3 <- update(nc.aov2,.~nc.int)
HSD.test(nc.aov3, "nc.int", group=TRUE, console=TRUE)

with(n2o.cum[!(n2o.cum$exp=="exp1"),], list(round(tapply(nc, interaction(Lt,soil), mean),0),
                   round(tapply(nc, interaction(Lt,soil), se),0)))
```

Analysis
```{r}
# data = co2.cum

co2.index$Lt <- revalue(co2.index$treat, c(Lt=1, int=1, Fc=0, C=0))
co2.index$Fc <- revalue(co2.index$treat, c(Lt=0, int=1, Fc=1, C=0))

co2.cum <- as.data.frame(cbind(co2.index[c(1:56),][-c(8,22),], co2.cum))
colnames(co2.cum)[21] <- "cc"
colnames(co2.cum)[4] <- "exp"

co2.cum <- droplevels(co2.cum)
co2.cum.org <- co2.cum
#co2.cum <- co2.cum.org

co2.cum <- co2.cum[-1,]
co2.cum <- co2.cum[-1,]


### Experiment 1

# multimodel averaging
x <- c("Lt","Fc","exp","soil")
cc.0 <- lm(cc~Lt*Fc*soil, co2.cum[!(co2.cum$exp=="exp2"),])
#t(combn(x,3))
cc.1 <- lm(cc~Lt*Fc, co2.cum[!(co2.cum$exp=="exp2"),])
cc.2 <- lm(cc~Lt*soil, co2.cum[!(co2.cum$exp=="exp2"),])
cc.3 <- lm(cc~Fc*soil, co2.cum[!(co2.cum$exp=="exp2"),])
#t(combn(x,2))
cc.4 <- lm(cc~Lt, co2.cum[!(co2.cum$exp=="exp2"),])
cc.5 <- lm(cc~Fc, co2.cum[!(co2.cum$exp=="exp2"),])
cc.6 <- lm(cc~soil, co2.cum[!(co2.cum$exp=="exp2"),])

cc.7 <- lm(cc~1, co2.cum[!(co2.cum$exp=="exp2"),])

AICctab(cc.0,  cc.1,  cc.2,  cc.3,  cc.4,  cc.5,  cc.6,  cc.7)

#as.character(rep("cc",c(1:15)))
#x <- factor(paste("cc",".",c(0:15),",",sep=""), levels=)
#print(x)
#x <- factor(paste("cc",".",c(0:15)," <- ",sep=""), levels=)
#as.data.frame(x)

summary.aov(cc.2)
summary.aov(cc.0)
par(mfrow=c(2,2))
plot(cc.0)
par(mfrow=c(1,1))
boxcox(cc.0, lambda=c(-50,50,5))
locator(1)

cc.0 <- update(cc.0,I(cc*10^-3)~.)
summary.aov(cc.0)
par(mfrow=c(2,2))
plot(cc.0)

cc.aov2 <- aov(cc~Lt*soil*Fc, co2.cum[!(co2.cum$exp=="exp2"),])
cc.aov2 <- update(cc.aov2,I(cc*10^-3)~.)
TukeyHSD(cc.aov2)
cc.int <- with(co2.cum[!(co2.cum$exp=="exp2"),], interaction(soil,treat)) # Double interaction factor
cc.aov3 <- update(cc.aov2,.~cc.int)
HSD.test(cc.aov3, "cc.int", group=TRUE, console=TRUE)

txt.cc = expression(paste("Carbon Dioxide", CO[2]," [mg ",CO[2],"-C ",kg^-1,"]"))
ggplot(co2.cum[!(co2.cum$exp=="exp2"),], aes(x=treat, y=cc), family="Arial") + 
  stat_boxplot(geom="errorbar", coef=1.5, lwd=0.2) + 
  geom_boxplot(fill="grey", lwd=0.2, outlier.size=0.4, outlier.shape=21) +  
  facet_grid( .~ soil , scales="free", space="free",labeller=label_parsed) + 
  ylab(txt.cc) + 
  xlab("L. terrestris") + 
  #scale_x_discrete(labels=c("absent", "present")) +
  ggtitle("Experiment 1") +
  mytheme


cc.int <- with(co2.cum[!(co2.cum$exp=="exp2"),], interaction(soil,Lt)) # Double interaction factor
cc.aov3 <- update(cc.aov2,.~cc.int)
HSD.test(cc.aov3, "cc.int", group=TRUE, console=TRUE)

with(co2.cum[!(co2.cum$exp=="exp2"),], list(round(tapply(cc, interaction(treat,soil), mean),0),
                                            round(tapply(cc, interaction(treat,soil), se),0)))
                                            
                                            
with(co2.cum[!(co2.cum$exp=="exp2"),], list(round(tapply(cc, interaction(Lt,soil), mean),0),
                   round(tapply(cc, interaction(Lt,soil), se),0)))


### Experiment 1

# multimodel averaging
x <- c("Lt","Fc","exp","soil")
cc.0 <- lm(cc~Lt*Fc*soil, co2.cum[!(co2.cum$exp=="exp1"),])
#t(combn(x,3))
cc.1 <- lm(cc~Lt*Fc, co2.cum[!(co2.cum$exp=="exp1"),])
cc.2 <- lm(cc~Lt*soil, co2.cum[!(co2.cum$exp=="exp1"),])
cc.3 <- lm(cc~Fc*soil, co2.cum[!(co2.cum$exp=="exp1"),])
#t(combn(x,2))
cc.4 <- lm(cc~Lt, co2.cum[!(co2.cum$exp=="exp1"),])
cc.5 <- lm(cc~Fc, co2.cum[!(co2.cum$exp=="exp1"),])
cc.6 <- lm(cc~soil, co2.cum[!(co2.cum$exp=="exp1"),])

cc.7 <- lm(cc~1, co2.cum[!(co2.cum$exp=="exp1"),])

AICctab(cc.0,  cc.1,  cc.2,  cc.3,  cc.4,  cc.5,  cc.6,  cc.7)

#as.character(rep("cc",c(1:15)))
#x <- factor(paste("cc",".",c(0:15),",",sep=""), levels=)
#print(x)
#x <- factor(paste("cc",".",c(0:15)," <- ",sep=""), levels=)
#as.data.frame(x)

summary.aov(cc.2)
par(mfrow=c(2,2))
plot(cc.2)
par(mfrow=c(1,1))
boxcox(cc.2)
# locator(1)
fligner.test(cc~interaction(Lt,soil), co2.cum[!(co2.cum$exp=="exp1"),])


cc.aov2 <- aov(cc~Lt*soil, co2.cum[!(co2.cum$exp=="exp1"),])
TukeyHSD(cc.aov2)
cc.int <- with(co2.cum[!(co2.cum$exp=="exp1"),], interaction(soil,Lt)) # Double interaction factor
cc.aov3 <- update(cc.aov2,.~cc.int)
HSD.test(cc.aov3, "cc.int", group=TRUE, console=TRUE)

txt.cc = expression(paste("Carbon Dioxide", CO[2]," [mg ",CO[2],"-C ",kg^-1,"]"))
ggplot(co2.cum[!(co2.cum$exp=="exp1"),], aes(x=Lt, y=cc), family="Arial") + 
  stat_boxplot(geom="errorbar", coef=1.5, lwd=0.2) + 
  geom_boxplot(fill="grey", lwd=0.2, outlier.size=0.4, outlier.shape=21) +  
  facet_grid( .~ soil , scales="free", space="free",labeller=label_parsed) + 
  ylab(txt.cc) + 
  xlab("L. terrestris") + 
  scale_x_discrete(labels=c("absent", "present")) +
  ggtitle("Experiment 2") +
  mytheme


cc.int <- with(co2.cum[!(co2.cum$exp=="exp1"),], interaction(soil,Lt,Fc)) # Double interaction factor
cc.aov3 <- update(cc.aov2,.~cc.int)
HSD.test(cc.aov3, "cc.int", group=TRUE, console=TRUE)

with(co2.cum[!(co2.cum$exp=="exp1"),], list(round(tapply(cc, interaction(Lt,soil), mean),0),
                   round(tapply(cc, interaction(Lt,soil), se),0)))
```

#### Multipanel plots for Gas fluxes of N2O and CO2
```{r Gas flux Multipanel plot}

measure=rep(c("mean","se"),each=16)

A=cbind(melt(cbind(measure=measure,rbind(n2o.cum.mean,n2o.cum.se)), id.vars= c(1:4))[,1:5],
      n2o.cum=melt(cbind(measure=measure,rbind(n2o.cum.mean,n2o.cum.se)), id.vars= c(1:4))$value,
      co2.cum=melt(cbind(measure=measure,rbind(co2.cum.mean,co2.cum.se)), id.vars= c(1:4))$value)
B=cbind(melt(cbind(measure=measure,rbind(n2o.means,n2o.se)),         id.vars= c(1:4))[,1:5],
      n2o=melt(cbind(measure=measure,rbind(n2o.means,n2o.se)),     id.vars= c(1:4))$value,
      co2=melt(cbind(measure=measure,rbind(co2.means,co2.se)),    id.vars= c(1:4))$value)

C=A[1:32,]
C[,6:7]=NA
C[,5]=1
A=rbind(C,A)
fig2.extended=cbind(A,B[6:7])
colnames(fig2.extended)[[5]] <- "day"
fig2.extended$day <- as.integer(fig2.extended$day)
fig2.x  <- melt(fig2.extended, id.vars= c(1:5)) 
fig2.x  <- dcast(fig2.x, treat+soil+variable+day+experiment~measure) 
fig2.x  <- fig2.x[order(fig2.x$experiment ),]
fig2.x$variable <- revalue(fig2.x$variable, c(n2o="paste(N[2],O)~Production~Rate",co2= "paste(CO[2])~Production~Rate", n2o.cum="Cumulative~paste(N[2],O)", co2.cum="Cumulative~CO[2]"))
fig2.x$variable <- factor(fig2.x$variable, levels=c("paste(N[2],O)~Production~Rate","Cumulative~paste(N[2],O)", "paste(CO[2])~Production~Rate","Cumulative~CO[2]"))

txt.ylab = expression(paste(CO[2],"[?g ",CO[2],"-C ",kg^-1,"]",
                                 "       ", 
                                 CO[2],"[?g ",CO[2],"-C ",kg^-1, h^-1,"]",
                                 "      ",
                                 N[2],"O [?g ",N[2],"O-N ",kg^-1,"]",
                                 "       ", 
                                 N[2],"O [?g ",N[2],"O-N ",kg^-1, h^-1,"]"))

pd = position_dodge(0.2)
ggplot(data=fig2.x, aes(x=day, y=mean, group=treat, shape=treat)) + 
  geom_point(size=1.5, position=pd, fill="white") + 
  geom_line (size=0.35, position=pd) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=4, size=0.2, position=pd) + 
  xlab("Days [d]") +
  ylab(txt.ylab) +
  facet_grid(variable~experiment+soil, scales ="free", labeller=label_parsed) + 
  scale_shape_manual(values=c(17,15,16,21), 
                     labels=c("L. terrestris","Interaction","F. candida", "Control")) + 
                     #limits=c("L. terrestris","F. candida","Interaction","Control")) +
  labs(shape="Treatment:") +
  mytheme + theme(legend.position="bottom")
````


#### Figure1 Multipanel plot with distinct variables
```{r Figure 1}
fig2  <- melt(fig2, id.vars= c(1:5)) 
fig2  <- dcast(fig2, treat+soil+variable+day+Experiment~measure) 
fig2$variable <- revalue(fig2$variable, c(nitox="paste(N[2],O)~Production~Rate", nitox_cum="Cumulative~paste(N[2],O)", carbon="Cumulative~CO[2]"))
fig2$variable <- factor(fig2$variable, levels=c("paste(N[2],O)~Production~Rate","Cumulative~paste(N[2],O)", "Cumulative~CO[2]"))

txt.cum_nitox = expression(paste(CO[2],"[?g ",CO[2],"-C ",kg^-1,"]",
                                 "                       ",
                                 N[2],"O [?g ",N[2],"O-N ",kg^-1,"]",
                                 "                       ",
                                 N[2],"O [?g ",N[2],"O-N ",kg^-1, h^-1,"]"))

pd = position_dodge(0.2)
ggplot(data=fig2, aes(x=day, y=mean, group=treat, shape=treat)) + 
  geom_point(size=2.0, position=pd, fill="white") + 
  geom_line (size=0.35, position=pd) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=4, size=0.2, position=pd) + 
  xlab("Days [d]") +
  ylab(txt.cum_nitox) +
  facet_grid(variable~soil, scales ="free", labeller=label_parsed) + 
  scale_shape_manual(values=c(21,16,15,17), 
                     labels=c("Control","F. candida","Interaction", "L. terrestris")) + 
  #                   limits=c("L. terrestris","F. candida","Interaction","Control")) +
  labs(shape=" ") +
  mytheme +
  theme(legend.position=c(0.57,0.9) )#+
       #legend.position="bottom"))

## ggsave("Figure 1.pdf", width=19, height=15, units="cm", useDingbats=FALSE)
```
 Appendix A
```{r Appendix A}
measure=rep(c("mean","se"),each=16)

A=cbind(melt(cbind(measure=measure,rbind(n2o.cum.mean,n2o.cum.se)), id.vars= c(1:4))[,1:5],
      n2o.cum=melt(cbind(measure=measure,rbind(n2o.cum.mean,n2o.cum.se)), id.vars= c(1:4))$value,
      co2.cum=melt(cbind(measure=measure,rbind(co2.cum.mean,co2.cum.se)), id.vars= c(1:4))$value)
B=cbind(melt(cbind(measure=measure,rbind(n2o.means,n2o.se)),         id.vars= c(1:4))[,1:5],
      n2o=melt(cbind(measure=measure,rbind(n2o.means,n2o.se)),     id.vars= c(1:4))$value,
      co2=melt(cbind(measure=measure,rbind(co2.means,co2.se)),    id.vars= c(1:4))$value)

C=A[1:32,]
C[,6:7]=NA
C[,5]=1
A=rbind(C,A)
fig2.extended=cbind(A,B[6:7])
colnames(fig2.extended)[[5]] <- "day"
fig2.extended$day <- as.integer(fig2.extended$day)
fig2.x  <- melt(fig2.extended, id.vars= c(1:5)) 
fig2.x  <- dcast(fig2.x, treat+soil+variable+day+experiment~measure) 
fig2.x  <- fig2.x[order(fig2.x$experiment ),]
fig2.x$variable <- revalue(fig2.x$variable, c(n2o="paste(N[2],O)~Production~Rate",co2= "paste(CO[2])~Production~Rate", n2o.cum="Cumulative~paste(N[2],O)", co2.cum="Cumulative~CO[2]"))
fig2.x$variable <- factor(fig2.x$variable, levels=c("paste(N[2],O)~Production~Rate","Cumulative~paste(N[2],O)", "paste(CO[2])~Production~Rate","Cumulative~CO[2]"))

txt.ylab = expression(paste(CO[2],"[?g ",CO[2],"-C ",kg^-1,"]",
                                 "                      ",
                                 #CO[2],"[?g ",CO[2],"-C ",kg^-1, h^-1,"]",
                                 #"                   ",
                                 N[2],"O [?g ",N[2],"O-N ",kg^-1,"]",
                                 "                      ",
                                 N[2],"O [?g ",N[2],"O-N ",kg^-1, h^-1,"]"))

pd = position_dodge(0.2)
ggplot(data=fig2.x[(fig2.x$experiment=="exp2" & !fig2.x$variable=="paste(CO[2])~Production~Rate"),], aes(x=day, y=mean, group=treat, shape=treat)) + 
  geom_point(size=2.0, position=pd, fill="white") + 
  geom_line (size=0.35, position=pd) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=4, size=0.2, position=pd) + 
  xlab("Days [d]") +
  ylab(txt.ylab) +
  facet_grid(variable~soil, scales ="free", labeller=label_parsed) + 
  scale_shape_manual(values=c(17,16,15,21), 
                     labels=c("L. terrestris","F. candida", "Interaction","Control")) + 
                     #limits=c("L. terrestris","F. candida","Interaction","Control")) +
  labs(shape=" ") +
  mytheme + theme(legend.position=c(0.57,0.9))

## ggsave("Appendix A.pdf", width=19, height=15, units="cm", useDingbats=FALSE)
```

  
  
# Stable Isotopes 15N Gas 
_______________________________________________________________________________________________________________

# Extrapolation of product ratio
```{r product ratio extrapolation}
str(gas15)
gas15$day <- as.numeric(gas15$day)

# For all together
pr.glm <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15)
summary(pr.glm)

par(mfrow=c(1,1))
with(gas15, plot(day, product.ratio))
     lines(gas15$day, pr.glm$fitted, type="l", col="red")
day=c(1,3,5,7,11,15,18,21,25,28,32,35,39,42,48)
day <- data.frame(day=day)
 x <- data.frame(predict(pr.glm, newdata = day , type = "response"))
str(day)

plot(cbind(day,x))

# for treat x soil
pr.glm1 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="Lt" & gas15$soil=="Loam"),])

pr.glm2 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="Fc" & gas15$soil=="Loam"),])

pr.glm3 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="LF" & gas15$soil=="Loam"),])

pr.glm4 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="C" & gas15$soil=="Loam"),])

pr.glm5 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="Lt" & gas15$soil=="Sand"),])

pr.glm6 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="Fc" & gas15$soil=="Sand"),])

pr.glm7 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="LF" & gas15$soil=="Sand"),])

pr.glm8 <- glm(cbind(nitox,nitox+dinitrogen) ~ day, family=binomial(logit),gas15[(gas15$treat=="C" & gas15$soil=="Sand"),])

pr.expol <- matrix(0,15,8)
pr.expol[,1] <- predict(pr.glm1, newdata=day,type = "response")
pr.expol[,2] <- predict(pr.glm2, newdata=day,type = "response")
pr.expol[,3] <- predict(pr.glm3, newdata=day,type = "response")
pr.expol[,4] <- predict(pr.glm4, newdata=day,type = "response")
pr.expol[,5] <- predict(pr.glm5, newdata=day,type = "response")
pr.expol[,6] <- predict(pr.glm6, newdata=day,type = "response")
pr.expol[,7] <- predict(pr.glm7, newdata=day,type = "response")
pr.expol[,8] <- predict(pr.glm8, newdata=day,type = "response")

colnames(pr.expol) <- c("Lt.Loam", "Fc.Loam", "Int.Loam", "C.Loam","Lt.Sand", "Fc.Sand", "Int.Sand", "C.Sand")


pr.expol <- data.frame(cbind(day,pr.expol))
pr.expol <- melt(pr.expol, id.vars=1)

pr.expol <-  concat.split.multiple(pr.expol, "variable", ".")

ggplot(pr.expol, aes(x=day,y=value)) + 
  geom_point(aes(colour=variable_1)) +
  facet_grid(~variable_2) + mytheme

colnames(n2o.means)[c(4:18)] <- c(1,3,5,7,11,15,18,21,25,28,32,35,39,42,48)

n2o.means.expol <- n2o.means[(n2o.means$experiment=="exp1"),]
n2o.means.expol <- melt(n2o.means.expol, id.vars=1:3)
n2o.means.expol$treat <- factor(n2o.means.expol$treat, levels=c("Lt","Fc","int","C"))
n2o.means.expol <- n2o.means.expol[with(n2o.means.expol, order(soil,treat)), ]

n2o.means.expol <- cbind(n2o.means.expol,pr.expol=pr.expol$value)
colnames(n2o.means.expol)[5] <- "nitox"
n2o.means.expol$total.x <- with(n2o.means.expol, nitox/pr.expol)

total.expol <- n2o.means.expol[,c(1,2,4,7)]
total.expol <- dcast(total.expol, treat+soil~variable)
total.expol <- total.expol[with(total.expol, order(soil,treat)),]
str(total.expol)

code <- total.expol[,c(1,2)]
total.expol <- total.expol[,-c(1,2)]

days=c(1,3,5,7,11,15,18,21,25,28,32,35,39,42,48)

ttx.rz=rbind(days,total.expol)

#ttx between two sampling dates
ttx.btw=matrix(rep(0,112),8)
for (i in 2:9)
{for (j in 1:14)
{Zeit=(ttx.rz[1,(j+1)]-ttx.rz[1,j])*24  #*24da Zeit d in std umgerechnet werden muss
 a=Zeit*ttx.rz[i,j]+0.5*Zeit*(ttx.rz[i,(j+1)]-ttx.rz[i,j])
 ttx.btw[(i-1),j]=a
}
}

# cumulative ttx for all samples
ttx.cum=matrix(rep(0,112),8)
for (i in 1:8) {
  for (j in 1:14) {
    a=sum(ttx.btw[i,1:j])
    ttx.cum[i,j]=a    
  }  
}

colnames(ttx.cum)<-days[-1]
ttx.cum <- cbind(code,ttx.cum)
ttx.cum <- melt(ttx.cum, id.vars=1:2)

ttx.cum2 <- subset(ttx.cum, !(treat %in% c("C","Fc") & soil %in% "Loam"))
pd = position_dodge(0.2)
ggplot(ttx.cum2, aes(x=variable, y=value, group=treat, shape=treat)) +
  facet_grid(~soil, scales="free", labeller=label_parsed) +
  geom_line(size=0.3, position=pd) +
  geom_point(size=1.5, position=pd, colour="black", fill="white") +
  scale_shape_manual(values=c(17,16,15,21), labels=c("L. terrestris", "F. candida","Interaction","Control")) +
  #scale_colour_manual(values=trt_pal) +
  #scale_colour_grey(start=0, end=0.6) +
  labs(shape="") +
  xlab("Days [d]") +
  ylab("total N extrapolated") +
  mytheme + theme(legend.position="bottom")  


ttx.pr <- n2o.means.expol[,c(1,2,3,4,5,6)]
ttx.pr <- ttx.pr[with(ttx.pr, order(variable,soil,treat)),]

ttx.cum <- rbind(cbind(ttx.pr[1:8,c(1,2)], variable=ttx.pr[1:8,4],value=rep(NA,8)),ttx.cum)
ttx.cum <- cbind(ttx.cum,nitox=ttx.pr$nitox, pr.expol=ttx.pr$pr.expol)
ttx.cum <- melt(ttx.cum, id.vars=1:3)
colnames(ttx.cum) <- c("treat", "soil", "day", "variable", "value")
ttx.cum$variable <- revalue(ttx.cum$variable, c(value="ttx") )
ttx.cum$variable <- factor(ttx.cum$variable, levels=c("nitox","pr.expol","ttx") )

pd = position_dodge(0.2)
ggplot(ttx.cum, aes(x=day, y=value, group=treat, shape=treat)) +
  facet_grid(variable~soil, scales="free", labeller=label_parsed) +
  geom_line(size=0.3, position=pd) +
  geom_point(size=1.5, position=pd, colour="black", fill="white") +
  scale_shape_manual(values=c(17,16,15,21), labels=c("L. terrestris", "F. candida","Interaction","Control")) +
  #scale_colour_manual(values=trt_pal) +
  #scale_colour_grey(start=0, end=0.6) +
  labs(shape="") +
  xlab("Days [d]") +
  ylab("total N extrapolated") +
  mytheme + theme(legend.position="bottom")  

ttx.cum2 <- subset(ttx.cum, !(treat %in% c("C","Fc") & soil %in% "Loam"))

pd = position_dodge(0.2)
ggplot(ttx.cum2, aes(x=day, y=value, group=treat, shape=treat)) +
  facet_grid(variable~soil, scales="free", labeller=label_parsed) +
  geom_line(size=0.3, position=pd) +
  geom_point(size=1.5, position=pd, colour="black", fill="white") +
  scale_shape_manual(values=c(17,16,15,21), labels=c("L. terrestris", "F. candida","Interaction","Control")) +
  #scale_colour_manual(values=trt_pal) +
  #scale_colour_grey(start=0, end=0.6) +
  labs(shape="") +
  xlab("Days [d]") +
  ylab("total N extrapolated") +
  mytheme + theme(legend.position="bottom")

#write.table(ttx.cum[(ttx.cum$day==48&ttx.cum$variable=="ttx"),],"ttx.cum.csv", sep=";")

```

wieviel N aus dem Pflanzenmaterial ging verloren?
wieviel h?her war N2 im vergleich zu n2o? Faktoren

# Isotopomers
```{r Isotopomers}

isom.data.600 <- with(isom.data.all, isom.data.all[N2O.conc > 600,])
isom.data.600$treat <- factor(isom.data.600$treat, levels=c("Lt", "Fc", "Int", "Control")) 

with(isom.data.all, plot(d18o,sp))
with(isom.data.600, plot(d18o,sp))
col=soil, fill=as.factor(Sampleset)),

ggplot(isom.data.600, aes(x=d18o, y=sp, shape=treat),  labeler=label_parsed) +
  #geom_point(aes(x=d18o, y=sp),size=1.6, lwd=0.01,col="black", shape=16, isom.data.600[(isom.data.600$soil=="Loam"),]) +
  geom_point(aes(x=d18o, y=sp),size=1.4, lwd=0.01,col="white", shape=16, isom.data.600[(isom.data.600$soil=="Loam"),]) +
  geom_point(size=2.0, lwd=0.2, col="black", show_guide=TRUE) +
  geom_point(size=1.2, lwd=0.2, aes(colour=as.factor(interaction(Sampleset,soil)))) +
  geom_rect( mapping=aes(xmin=10, xmax=20, ymin=-10, ymax=0), color="grey", alpha=0.01) +
  geom_rect( mapping=aes(xmin=40, xmax=50, ymin=33, ymax=36), color="grey", alpha=0.01) +
  geom_rect( mapping=aes(xmin=30, xmax=40, ymin=34, ymax=37), color="grey", alpha=0.01) +
  geom_abline(intercept = c(-12.5,0), slope = 0.5, alpha=0.1)+
  geom_abline(intercept = c(-20,0), slope = 0.5, alpha=0.1)+
  geom_abline(intercept = c(-5,0), slope = 0.5, alpha=0.1)+
  geom_abline(intercept = c(-8,0), slope = 0.2, alpha=0.1)+
  geom_abline(intercept = c(-14,0), slope = 0.2, alpha=0.1)+
  geom_abline(intercept = c(-2,0), slope = 0.2, alpha=0.1)+
  annotate("text",x=35, y=35.5, label="fungal nitrification", size=1.6, colour="black" ,face="italic") +
  annotate("text",x=45, y=34.5, label="nitrification", size=1.6, colour="black" ,face="italic") +
  annotate("text", x=15, y=-5, label="denitrification", size=1.6, col="black") +
  annotate("text", x=45, y=20, label="y=0.5x+c", size=1.6, col="black") +
  annotate("text", x=40, y=-8, label="y=0.2x+c", size=1.6, col="black") +
  ylab("Site preference [\211]") +
  xlab(expression(paste({delta}^18*O[soil]," [\211]"))) +
  scale_y_continuous(limits=c(-10,40),breaks=seq(-10,40,10)) +
  #scale_shape_manual(values=c(21,23,24,22)) +
  scale_shape_manual(labels=c("L. terrestris","F. candida","Interaction","Control"),values=c(17,16,15,18)) +
  scale_colour_manual("soil texture",labels=c("Loam (day 3)", "Loam (day 21)", "Sand (day 21)"),values=c("grey", "white", "black")) +
  mytheme + theme(legend.position="none", 
                  axis.title.y = element_text(size=8, face="plain", family="Times New Roman"))
  
+ theme(legend.position=c(0.1,0.75), 
                  legend.text=element_text(size=6,face="plain", family="Times New Roman"),
                  text=element_text(lineheight=0.4),
                  legend.margin=unit(0.2, "cm")) +
  guides(shape = guide_legend(label.theme =element_text(family="Times New Roman", face = "italic", size=6, angle=0)))


                  #ggsave("Figure 5.pdf", width=9, height=7, units="cm", useDingbats=FALSE)

locator(1)


with(isom.data.600, list(round(tapply(sp, list(treat, soil, Sampleset), mean),0),
                   round(tapply(sp, list(treat, soil, Sampleset), se),0)))



isom.lm1 <- lm(sp~interaction(Sampleset,soil),isom.data.600)
summary(isom.lm1)



isom.test  <-  cbind(isom.data.600$sp, isom.data.600$d18o)
isom.mv1 <- manova(isom.test~interaction(Sampleset,soil),isom.data.600)
summary(isom.mv1, test="Pillai")
summary.aov(isom.mv1)

isom.aov1 <- aov(sp~interaction(Sampleset,soil),isom.data.600)
summary(isom.aov1)
TukeyHSD(isom.aov1)

isom.aov2 <- aov(d18o~interaction(Sampleset,soil),isom.data.600)
summary(isom.aov2)
TukeyHSD(isom.aov2)


with(isom.data.600, by(sp, interaction(soil,Sampleset),summary))
with(isom.data.600, by(d18o, interaction(soil,Sampleset),summary))

# Reduction of n2o to n2, Rayleigh Equation
# Rayleigh <- SP = SP0 + eSP*ln(f) // Sp = Sp0 + eSP*(1-f)
# nach F aufgel?st:
# f = e^((SP-SP0)/eSP) // f = 1-(SP-SP0)/eSP

isom.loam.set2 <- isom.data.600[with(isom.data.600,(Sampleset==2 & soil=="Loam")),]
sp <- mean(isom.loam.set2$sp)+1000
sp0 <- -5+1000
fracfac <- 1000/-5

f  <-  (sp/sp0)^fracfac
round(f,2)

```


# Quantification of N-Loss
```{r N-Loss}

# N-Loss from n2o emissions (with soil derived!)

n2o.cum.mean[1:8,c(1,2,3,17)]

x <- n2o.cum.mean[1:8,17]

x <- x*10^-3

fac.ha*x*10^-6


# N Loss from 15N isotope sampling

x2 <- with(gas15.cum2.mean[gas15.cum2.mean$day==40,], as.numeric(dinitrogen.mean)+as.numeric(nitox.gc.mean))

N.loss15 <- cbind(Lt=factor(rep(c(1,0,1,0),2)), gas15.cum2.mean[gas15.cum2.mean$day==40,], N.loss=x2*10^-3)

total1 <- with(N.loss15, list( tapply(N.loss, list(Lt,soil), mean),
                        tapply(N.loss, list(Lt,soil), se)))
total2 <- with(N.loss15, list( tapply(N.loss, Lt, mean),
                        tapply(N.loss, Lt, se)))
total1

total2

N.input = 61.5 #[mg]
N.lost = c(1.52, 1.23,0.18) #[mg/kg]

#percentage N-loss
N.lost/N.input*100

# calculated N-loss ofr 1 ha
ha = 10000 #[m^2]
bulk = 1200 #[kg m^-3]
depth = 0.2 #[m]

fac.ha = ha*depth*bulk
fac.ha = 2400000 #[kg]
fac.ha*N.lost*10^-6 #[kg/ha]

```



