---
title: "V1 - Soil fauna effects on gaseous nitrogen"
author: "Quentin Schorpp"
date: "Monday, December 15, 2014"
output: html_document
---
e-mail: quentin.schorpp@ti.bund.de

Coauthors: *Catharina Riggers, Dominika Lewicka-Szczebak, Reinhard Well, Stefan Schrader*

This is the Analysis of a microcosm experiment performed at the Thünen Institute of Biodiversity in spring 2013 (May-July) in Braunschweig, Germany. The Experiment aimed to investigate the influence of soil fauna groups Lumbricidae (*L. terrestris*) and Collembola (*F. candida*) on gaseous nitrogen emissions. The data on emissions consists of three different qualities:

1. emissions of N[2]O and CO[2] were measured continously during the whole duration of the experiment in intervals lasting several days. That data is in units of ppb or ppm.  
2. at four dates the microcosms were incubated with artificial atmosphere and 15N isotopic signatures of N2O and N2 were measured.
3. at one particular date samples for isotopomere measurements were taken.  

But before analysing the Gas data, the analysis begins with data on soil fauna (growth and ernichment).  

In the framework of this experiment Catharina Riggers wrote her bachelor thesis.

1.Get started
===========
#### 1.1 Set working directory
```{r 1-1 working directory, echo=FALSE}
setwd("E:/Quentin_Schorpp_20.11.2014/Arbeitsprozess/V1_Analysis")
#setwd("D:/Quentin Schorpp/Arbeitsprozess/V1_Analysis")
#setwd(choose.dir())
```

#### 1.2 Load required packages
```{r 1-2 packages, warning=FALSE, message=FALSE, echo=FALSE}
library(ggplot2)
library(reshape2)
library(grid)
library(plyr)
library(MethComp)
library(agricolae)
library(foreign)
library(multcomp)
library(gridExtra)
library(extrafont)
library(effects)
library(MASS)
library(splitstackshape)
library(bbmle)
````

#### 1.3 Additional functions
```{r 1-3 Additional Functions, echo=FALSE}
# Function for Bland Altman Plot
#------------------------------------
baplot <- function(m1, m2, ...) {
        # m1 and m2 are the measurements
        means <- (m1 + m2) / 2
        diffs <- m1 - m2
        mdiff <- mean(diffs)
        sddiff <- sd(diffs)
        # Compute the figure limits
        ylimh <- mdiff + 3 * sddiff
        yliml <- mdiff - 3 * sddiff
        # Plot data
        plot(diffs ~ means, xlab = "Average values",
             ylab = "Differences", ylim = c(yliml, ylimh), ...)
        abline(h = mdiff) # Center line
        # Standard deviations lines
        abline(h = mdiff + 1.96 * sddiff, lty = 2)
        abline(h = mdiff - 1.96 * sddiff, lty = 2)
        }

# Function for layout
#-------------------------------------
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)

# Function for Standard error
#----------------------------
se <- function(x) sqrt(var(x)/length(x))
```

#### 1.4 ggplot2 - Theme
```{r 1-4 mytheme, echo=FALSE}
mytheme = 
        theme_bw() + 
        theme(strip.background = element_rect(color = "grey", fill="black", size=0.1),
              strip.text.x = element_text(size=8,  colour="white", face="italic"),
              strip.text.y = element_text(size=8,  colour="white", face="italic"),
              axis.text.x = element_text(size=7),
              axis.title.x = element_text(size=8,face="bold", family="Times New Roman"),
              axis.text.y = element_text(size=7),
              axis.title.y = element_text(size=8,face="bold", family="Times New Roman"),
              axis.line = element_line(size=0.25),
              axis.ticks = element_line(size=0.25),
              plot.title = element_text(size=11,face="bold", family="Times New Roman"),
              panel.margin = unit(0, "lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_rect(colour="black", size=0.2, fill=NA),
              legend.key=element_blank(),
              legend.background=element_blank(),
              legend.text=element_text(size=8,face="italic", family="Times New Roman"),
              legend.title=element_text(size=8))
```

#### 1.5 Load data
```{r 1-5Load data, echo=FALSE}
ew.bm <- read.delim("V1_worms_biomass.txt")
col.pop <- read.delim("V1_col_population.txt")
ew.iso <- read.delim("V1_worms_isotopes.txt")
col.iso <- read.delim("V1_col_isotopes.txt")
n2o.ppb <- read.delim("V1_N2Oppb.txt")
co2.ppm <- read.delim("V1_CO2ppm.txt")
gasflow <- read.delim("V1_Gasflow.txt")
fig2 <- read.delim("V1_Fig_2_data.txt")
gas15 <- read.delim("V1_15Ngas.txt")
isom.data.all <- read.delim("V1_Isotopomere.txt")
```

2. Soil Fauna
=============
## 2.1 Eathworms survival rate and biomass
Earthworm biomass was measured in g after earthworm guts were voided. Mesurements took place before and after the experiment.
During the Experiment four Individual of L. terrestris died in Microcosms with the interaction Treatment in the 15N experimental part / approach.

### 2.1.1 Survival Rates:
```{r Ew Survival rates, echo=FALSE}
Lt.Exp1  = 100*(1-4/(4*2*2*2)) # = 87.5 %
Lt.Exp2  = 100*(1-0/(3*2*2*2)) # = 100 %
Lt.total = 100*(1-4/((4*2*2*2)+(3*2*2*2))) # = 92.86 %
# Formula: 100*(1-dead worms/all worms) [%]
```
In total the survival rate of earthworms was `r Lt.total`%. Only in Experiment 1 (labelled leaf litter) 4 earthworms died. The dead earthworms were found in interaction treatments (MC8,MC22) in one microcosm with loam and one with sand, each. Therefore the survival rate for Exp1 was `r Lt.Exp1`% and two microcosms were lost for Analysis. In the biomass dataset these two microcosms were already deleted.

### 2.1.2a Earthworm biomass
```{r Ew-biomass, echo=FALSE, warning=FALSE}
# Data Exploration
str(ew.bm)
summary(ew.bm[,-1])

# Optimal binsizes
ew.bm.bs1 = diff(range(ew.bm$before))/26
ew.bm.bs2 = diff(range(ew.bm$after))/26
ew.bm.bs3 = diff(range(ew.bm$diff))/length(ew.bm$diff)

# Histograms Before/After
ew.bm.melt  <-  melt(ew.bm[,-c(7,8)], id.vars=c(1:4))
ew.bm.p1 <- ggplot(ew.bm.melt, aes(x=value)) +
                geom_histogram( binwidth = ew.bm.bs1, fill="lightblue3", colour="black") + 
                geom_line(stat="density") +
                xlim(6.5,11) + 
                xlab("Biomass [g]") +
                ylab("Frequency") +
                ggtitle("Histograms EW-biomass \n Before/After") +
                facet_grid(variable ~ .) + mytheme

# normally distributed diferences?
ew.bm.p2 <- ggplot(ew.bm, aes(x=diff)) +
                geom_histogram(binwidth = bs.worms3, fill="lightblue3", colour="black") + 
                geom_line(stat="density") + 
                xlab("differences [g]") + 
                ylab("Frequency") + 
                ggtitle("Histogram of differences\nnormal distribution?") + mytheme

# One figure in row 1 and two figures in row 2
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
        print(ew.bm.p1, vp = vplayout(1, 1))
        print(ew.bm.p2, vp = vplayout(1, 2))
```
  
### 2.1.2b Analysis
``` {r Analysis1, echo=FALSE, fig.width=8, fig.height=7}
shapiro.test(ew.bm$diff) 
# kann bei diesem Test eine Signifikanz (p < 0.05) festgestellt werden, so liegt keine Normalverteilung vor.


#ggplot with qqnorm and qqline
vec <- ew.bm$diff
y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
x <- qnorm(c(0.25, 0.75))
slope <- diff(y)/diff(x)
int <- y[1L] - slope * x[1L]
d <- data.frame(resids = vec)

ew.bm.p3 <- ggplot(d, aes(sample = resids)) + 
        stat_qq() + 
        geom_abline(slope = slope, intercept = int, col="red") + 
        ggtitle("qqplot") +
        mytheme

# Bland Altman plot
ew.bm$diff.mean <- (ew.bm$after+ew.bm$before)/2
df2 <- ddply(ew.bm,.(rep("a", length(diff))),summarise,mean=mean(diff, na.rm = TRUE),
             sd=sd(diff, na.rm = TRUE))

ew.bm.p4 <- ggplot(ew.bm, aes(diff.mean, diff)) + 
                geom_point(na.rm=TRUE) + 
                geom_hline(data=df2,aes(yintercept=c(round(mean,3),
                                                     round(mean+2*sd,3),
                                                     round(mean-2*sd,3))),
                           linetype=c(1,2,2), color='blue') +
                ggtitle("Bland Altman plot") +
                mytheme
# The Bland 
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
print(ew.bm.p3, vp = vplayout(1, 1))
print(ew.bm.p4, vp = vplayout(1, 2))


# ew.bm.meth <- Meth(ew.bm[,c(5,6)], y=c("after","before")) # it was necessary, that no other variables from ew.bm are in the Meth object!!!
# BA.plot(ew.bm.meth, axlim=c(8,10), diflim=c(-3,1))
```
Paired T-Test
```{r, echo=FALSE}
# t.test
# with(ewfw, t.test(diff))
# with(ew.bm, t.test(before, after, paired=T, alternative="greater"))
ew.bm.ttest <- with(ew.bm, t.test(after, before, paired=T, alternative="less")) 
print(ew.bm.ttest)
# Defining alt="less", means check whether the mean of the values contained in the vector a is less of the mean of the values contained in the vector b (t.test(a,b,paired=TRUE, alt="less"))
````
ANOVA
``` {r, echo=FALSE} 
# Were Changes in L. terrestris Biomass affected by Treatments?
summary(aov(diff ~ treat*soil*exp, data=ew.bm))
# NO it was not.

ew.bm.values <- list(before=c(mean(ew.bm$before),"+/-", se(ew.bm$after)),after=c(mean(ew.bm$after),"+/-",se(ew.bm$before)))
print(ew.bm.values)
```
All earthwroms lost Biomass during the experiment(paired t-test;`r ew.bm.ttest[c(1,2,3)]`). Before the experiment earthworm weighed `r ew.bm.values$before` [g] and after `r ew.bm.values$after` [g]. The Biomass loss of earthwroms did not depend on any of the treatments.


##2.2 Collembolan Population growth
Collembolans grew very well during the experiment and reached densities hardly observed in agricultural systems.

### 2.2.1 Growth Factor
```{r Collembolan population}
col.pop.initial = 238 #Ind/MC
col.pop.growth = col.pop$surface/238
col.pop$growth_f = col.pop.growth

range(col.pop.growth); mean(col.pop.growth); se(col.pop.growth)
```
In average the collembolan populations grew 46 fold +/- 4. 

```{r, Boxplots, fig.width=7, fig.height=6, echo=FALSE}
# Boxplot Graphical Data Exploration
col.pop$exp <- revalue(col.pop$exp, c(Exp1="Experiment~1",Exp2="Experiment~2"))
col.pop$treat <- revalue(col.pop$treat, c(C ="F.~candida",RC="Interaction"))

ggplot(col.pop, aes(x= treat, y=surface),labeller=label_parsed) +
        stat_boxplot(geom="errorbar", coef=1.5, lwd=0.2) + 
        geom_boxplot(fill="light blue", lwd=0.2, outlier.size=0.4, outlier.shape=21) +
        ylab(expression(paste("Collembolan density [Ind ",MC[surface]^-1,"]"))) +
        xlab("Treatment") +
        ggtitle(expression(paste("Boxplots for collembolan density per ",MC[surface]))) +
        facet_grid(exp~soil, labeller=label_parsed) + 
        scale_x_discrete("Treatment", labels=expression(italic(F.~candida), Interaction)) +
        mytheme
````

In 3 out of 4 cases Presence of Lumbricus terrestris led to increased Collembolan population growth. The only exception is the loamy soil treatment in Exp1. Here the opposite was observed.
Furthermore loamy soil treatments increased collembolan population growth in both experiments compared to sandy soil (`r tapply(col.pop$surface, col.pop$soil,mean)`, Loam, Sand).
Hypothesis: Low Nitrogen content of food material leads to decreased collembolan populations in loamy soil if L. terrestris is present.

### Analysis of Collembolan populations
```{r Analysis2}
# Table of means
with(col.pop, tapply(surface, list(soil, treat, exp), mean))

# Analysis of variance
cp.lm1 <- lm(surface ~ treat*soil*exp, data=col.pop);cp.lm1; summary.aov(cp.lm1);summary.lm(cp.lm1)
cp.lm2 <- lm(surface ~ treat*soil*exp-exp:treat:soil, data=col.pop);cp.lm2; summary.aov(cp.lm2);summary.lm(cp.lm2)
# Soil and Experiment are significant in the main effects
# Treatments is significant in interaction with soil and with experiment

# The coefficients table of summary.lm has as many rows, as there are parameters in the model.The Intercept is the only mean value in the table. The second column contains the unreliability estimates. The firs ROW contains the standard error of a MEAN. The Other four rows contain the standard errors of  the DIFFERENCE between two means. The p-values are misleading, suggesting, wrongly, that there are four significant contrasts for this model.
# All significant differences are differences to Exp1-Loam-F.candida

# Effect Sizes of Main Effects
plot.design(surface ~ soil*exp*treat-exp:treat:soil, data=col.pop)

# The populations were larger in loamy soil
data.frame(with(col.pop, list(mean=tapply(surface, soil, mean),
                              SE=tapply(surface, soil, se))))

# The populations were larger in Experiment 2
data.frame(with(col.pop, list(mean=tapply(surface, exp, mean),
                              SE=tapply(surface, exp, se))))

cp.aov <- aov(log(surface) ~ soil*exp*treat, data=col.pop)
# ANOVA plots
par(mfrow=c(2,2))
par(mar=c(3,4,1.5,2), mgp=c(1.5,0.5,0))
plot(cp.aov)

TukeyHSD(cp.aov)
# Unterschiede zw. den treatments innerhalb desselben Experiments und derselben soil texture sind ALLE nicht significant!!!  

# More Multiple Comparisons:
# Variant 1:
col.pop.int <- cp.int <- with(col.pop, interaction(exp,soil,treat)) # Triple interaction factor
col.aov2 <- aov(surface ~ cp.int, data=col.pop)
HSD.test(col.aov2, "cp.int", group=TRUE, console=TRUE)
#Variant 2:  
col.tuk1 <- glht(col.aov2, linfct = mcp(cp.int = "Tukey"))
summary(col.tuk1)          # standard display
col.tuk.cld <- cld(col.tuk1)   # letter-based display
par(mai=c(1,1,1.1,0.2),
    mfrow=c(1,1),
    mgp=c(2,1,0),
    cex=0.8) # mai specifiec margin size in inches
plot(col.tuk.cld, cex=0.5, las=2, col="grey", xaxt="n")
axis(1, at=cp.int,labels=FALSE)
text(cp.int, labels=cp.int, par("usr")[3], adj=c(1.2,1.2), xpd=TRUE, srt=45, cex=0.8)


# Effect sizes of interactions
col.pop.effects <- allEffects(cp.lm2)
plot(col.pop.effects, "treat:soil")
plot(col.pop.effects, "treat:exp")
plot(col.pop.effects, "soil:exp")

model.tables(cp.aov, "means", se=TRUE)
with(col.pop, tapply(surface, list(treat,soil), mean))
with(col.pop, tapply(surface, list(treat,soil), se))

with(col.pop, tapply(surface, list(treat,exp), mean))
with(col.pop, tapply(surface, list(treat,exp), se))

with(col.pop, tapply(surface, list(soil,exp), mean))
with(col.pop, tapply(surface, list(soil,exp), se))

col.pop.effects <- allEffects(cp.lm1)
plot(col.pop.effects, "treat:soil:exp")

# Interessieren würde mich wie viel größer die Populationen im Schnitt in Anwesenheit von L. terrestris waren.
# Datensatz zu differenzen zw. Interaktion und Single-Species Treatments
# PROBLEM: Welche Mikrokosmen bilden Paare? 
# Bilden aller paarweise DIfferenzen und/oder nur für selbe Faktoren Kombination aus Exp und soil?
# Wie subtrahiert man Standardfehler voneinander (gibt T-Test Aufschlüsse?)

# Mein reduzierter Versuch:
treat.sand.exp1 = a = 9629 - 6319
treat.sand.exp2 = b = 18781 - 14766
treat.loam.exp2 = c = 11246 - 4487
cp.diff.means = c(a,b,c)
mean(cp.diff.means); se(cp.diff.means)

```

### Tukey HSD plots Two Way Interactions (optional)
```{r More Tukey HSD, echo=FALSE, eval=FALSE}
cp.int <- with(col.pop, interaction(treat, soil))
aov.col1 <- aov(surface ~ cp.int, data=col.pop)
HSD.test(aov.col1, "cp.int", group=TRUE, console=TRUE)
tuk <- glht(aov.col1, linfct = mcp(cp.int = "Tukey"))
summary(tuk)          # standard display
tuk.cld <- cld(tuk)   # letter-based display
opar <- par(mai=c(1,1,1.5,1))
plot(tuk.cld, las=2)
par(opar)

cp.int <- with(col.pop, interaction(exp, soil))
aov.col1 <- aov(surface ~ cp.int, data=col.pop)
HSD.test(aov.col1, "cp.int", group=TRUE, console=TRUE)
tuk <- glht(aov.col1, linfct = mcp(cp.int = "Tukey"))
summary(tuk)          # standard display
tuk.cld <- cld(tuk)   # letter-based display
opar <- par(mai=c(1,1,1.5,1))
plot(tuk.cld)
par(opar)

cp.int <- with(col.pop, interaction(exp, treat))
aov.col1 <- aov(surface ~ cp.int, data=col.pop)
HSD.test(aov.col1, "cp.int", group=TRUE, console=TRUE)
tuk <- glht(aov.col1, linfct = mcp(cp.int = "Tukey"))
summary(tuk)          # standard display
tuk.cld <- cld(tuk)   # letter-based display
opar <- par(mai=c(1,1,1.5,1))
plot(tuk.cld)
par(opar)
```

# Analysis of 15N enrichment in animal tissue
## 15N enrichment in L. terrestris
```{r 15N L. terrestris}
bs4 <- diff(range(ew.iso$atpercent))/length(ew.iso$atpercent)
ggplot(ew.iso, aes(x=atpercent)) + 
        geom_histogram(binwidth = bs4 , fill="lightblue3", colour="black") + 
        xlab(expression(paste(.^15,"N atom%"))) + 
        ylab("frequency") +
        ggtitle(expression(paste("Histogram of", .^15,"N Enrichment in tissue of ",italic(L.~terrestris))))


# Overlap between Loam and Sand in %
100*((with(ew.iso[ew.iso$soil=="S",], max(atpercent)) - with(ew.iso[ew.iso$soil=="L",], min(atpercent)))/with(ew.iso,max(atpercent) - min(atpercent)))

maxRows <- by(ew.iso, ew.iso$soil, function(X) X[which.max(X$atpercent),])
minRows <- by(ew.iso, ew.iso$soil, function(X) X[which.min(X$atpercent),])
do.call("rbind", maxRows)
do.call("rbind", minRows)

ew.iso.plot <-  qplot(soil, atpercent, colour=treat, data=ew.iso) + 
        geom_point(size=2.5) +
        annotate("text", x=2,y=1.01+.02, label="max=1.01", vjust=0, col="red", size=4) +
        annotate("text", x=1,y=0.74+.02, label="min=0.74", vjust=0,col="red", size=4) +
        xlab("Soiltype") + 
        ylab(expression(paste(.^15,"N atom%",sep=""))) + 
        ggtitle(expression(paste("Label in ",italic("L. terrestris"))))+
        scale_x_discrete(breaks=c("L","S"), labels=c("Loam", "Sand"))+
        labs(colour="Treatment")+
        scale_colour_discrete(labels=c("L.terrestris", "L. terrestris + F. candida")) + 
        theme(panel.background= element_rect(fill="white"),
              panel.grid.major = element_line(colour="grey"),
              panel.grid.minor = element_line(colour="grey", linetype="dashed", size=0.2),
              panel.border= element_rect(colour="black", fill=NA, size=1),
              axis.text = element_text(colour="black", size=rel(1.3)),
              axis.text.x = element_text(face="italic"),
              axis.title=element_text(size=rel(1.4)),
              legend.text=element_text(face="italic"))
ew.iso.plot
````

15N Enrichment in animal tissue
===============================
## 3. Analysis of 15N enrichment in L. terrestris
```{r Analysis3}

M.ew.iso <- lm(atpercent~treat*soil, data=ew.iso);anova(M.ew.iso);summary(M.ew.iso)
M.ew.iso <- lm(atpercent~soil, data=ew.iso); anova(M.ew.iso);summary(M.ew.iso)
lm.influence(M.ew.iso)
influence.measures(M.ew.iso)$is.inf

par(mfrow=c(2,2),
    mar=c(3,3,3,3))
plot(M.ew.iso)

shapiro.test(ew.iso$atpercent)
fligner.test(atpercent~soil, data=ew.iso)

t.test(atpercent~soil, ew.iso)
t.test(atpercent~soil,ew.iso, alternative="greater")

with(ew.iso, list( tapply(atpercent, soil, mean),
                   tapply(atpercent, soil, se)))
```
L. terrestris was significantly more enriched in loamy soil, compared to sandy soil 15N in tissue was with 1.04 +/- 0.08 atom% greater than  0.75 +/- 0.08 in sandy soil.

## 15N enrichment in F.candida
```{r 15N F. candida}
#Histograms
par(mfrow=c(2,1))
hist(col.iso$atpercent)
hist(col.iso$atpercent, main="breaks=12", breaks=12) # Too narrow --> breaks in Histogram

list(range(col.iso$atpercent),
     range(col.iso$atpercent)[2] -
             range(col.iso$atpercent)[1])

#There's a differnce between the Minima and Maxima of 0.896 atom percent
# what differnce in atpercent is meaningful?

with(col.iso, list(median(atpercent),
                   mean(atpercent))) # median and mean are the same: 6.4 atom percent

icept = median(col.iso$atpercent)

col.iso.plot <- 
        qplot(soil, atpercent, colour=treat, data=col.iso) + 
        geom_point(size=2.5) + 
        geom_hline(yintercept=icept, linetype="dashed", colour="red") +
        annotate("text", label="Midline Max(Loam) - Min(Sand) y=6.4", x=1.5, y=6.44, size=4,colour="red") +
        ggtitle(expression(paste("Label in ",italic("F. candida"))))+  
        xlab("soiltexture") + 
        ylab(expression(paste(.^15,"N atom%",sep=""))) + 
        scale_x_discrete(breaks=c("L","S"), labels=c("Loam", "Sand"))+
        scale_colour_discrete(labels=c("F.candida", "F. candida \n + L. terrestris")) + 
        labs(colour="Treatment")+
        theme(panel.background= element_rect(fill="white"),
              panel.grid.major = element_line(colour="grey"),
              panel.grid.minor = element_line(colour="grey", linetype="dashed", size=0.2),
              panel.border= element_rect(colour="black", fill=NA, size=1),
              axis.text = element_text(colour="black", size=rel(1.3)),
              axis.text.x = element_text(face="italic"),
              axis.title=element_text(size=rel(1.4)),
              legend.text=element_text(face="italic"))
col.iso.plot
````

```{r Analysis4}
M.col.iso1 <- with(col.iso, aov(atpercent~treat*soil)); summary(M.col.iso1)
M.col.iso2 <- with(col.iso, aov(atpercent~soil)); summary(M.col.iso2)
# What other nutritional source could F. candida have had in Loam?

anova(M.col.iso1, M.col.iso2)
AIC(M.col.iso1, M.col.iso2)

M.col.iso3 <- lm(atpercent~soil, data=col.iso)
lm.influence(M.col.iso3)
influence.measures(M.col.iso3)$is.inf

t.test(atpercent~soil, col.iso, alternative="less")

with(col.iso, list( tapply(atpercent, soil, mean),
                    tapply(atpercent, soil, se)))

mean(col.iso$atpercent)
se(col.iso$atpercent)

```

```{r Figure 1}
organism <- factor(rep(c("italic(L.~terrestris)", "italic(F.~candida)"), each=12), levels=c("italic(L.~terrestris)", "italic(F.~candida)"))
ew.col.iso <- rbind(ew.iso, col.iso)
ew.col.iso$organism <- organism
ew.col.iso$soil <- revalue(ew.col.iso$soil, c(L="Loam", S="Sand"))

#long cut way to find number of facets
len <- length(levels(ew.col.iso$organism))# *  length(levels(mtcars$am))
vars <- data.frame(organism=levels(ew.col.iso$organism))
y = with(ew.col.iso, tapply(atpercent,organism,max)-0.025)
dat <- data.frame(x = rep(0.5, len), y = y, vars, labs=LETTERS[1:len])

ggplot(ew.col.iso, aes(x=soil, y=atpercent), family="Arial") + 
        stat_boxplot(geom="errorbar", coef=1.5, lwd=0.2) + 
        geom_boxplot(fill="grey", lwd=0.2, outlier.size=0.4, outlier.shape=21) +  
        facet_grid(organism ~ . , scales="free", space="free",labeller=label_parsed) + 
        ylab("atom %") + 
        xlab("soil texture") + 
        geom_text(aes(x, y, label=labs, group=NULL),data=dat) +
        # ggtitle(expression(paste(.^15,"N enrichment in animal tissue"))) +
        mytheme



#setwd("D:/Quentin Schorpp/Schreibtisch/2 - Laborversuche/Mikrokosmen/Manuskript/Entwurf")
#ggsave("Figure1.pdf", width=9, height=7, units="cm", useDingbats=FALSE) 
#embed_fonts("Figure1.pdf", outfile="Figure1.pdf")
#ggsave("Figure1.svg", width=9, height=7, units="cm") 
```


